### 2017-06-08 sheets/talk-6/sheet-6.tex
Sample solution Talk 6.8

    Lets reconcider the definition of the \emph{product}:
    For every object $A \in \cat{A}$ and arrows $f_1 \from A \to X, f_2 \from A \to Y$ we get that there exists a unique map
    $\bar{f} \from A \to P$, such that
    \[ \xymatrix{
        &A \ar[ldd]_{f_1} \ar@{.>}[d]|{\bar{f}\vphantom{\bar{\bar{f}}}}
        \ar[rdd]^{f_2}&       \\
                &P \ar[ld]^{p_1} \ar[rd]_{p_2}                  &       \\
        X       &                                               &Y
    } \]
    commutes.
    Now we can instanciate $A$ with $P'$, $f_1$ with $p_1$, $f_2$ with $p_2$. By analogous instanciations we get the commuting diagramm
    \[ \xymatrix{
      & P \ar@/_/[ldd]_{p_1} \ar@/^/[rdd]^{p_2} & \\
      & P \ar[ld]^{p_1} \ar@/_/@{.>}[dd]|{\bar{f}} \ar@{.>}[u]|{f} \ar[rd]_{p_2} & \\
      X & & Y \\
      & P' \ar[lu]_{p_1'} \ar@/_/@{.>}[uu]|{\tilde{f}} \ar@{.>}[d]|{f'} \ar[ru]^{p'_2} & \\
      & P' \ar@/^/[luu]^{p'_1} \ar@/_/[ruu]_{p'_2} &
    } \]
    for exactly one $f, f', \tilde{f}, \bar{f}$. With the inner triangles we get:
    \begin{alignat*}{2}
      p_1  \of \tilde{f} &= p'_1 &\qquad p_2  \of \tilde{f} &= p'_2 \\
      p'_1 \of \bar{f}   &= p_1  &       p'_2 \of \bar{f}   &= p_2
    \end{alignat*}
    By substitution we get
    \begin{alignat*}{2}
      p_1   \of (\tilde{f} \of \bar{f}) &= p_1  &\qquad p_2  \of (\tilde{f} \of \bar{f}) &= p_2 \\
      p'_1  \of (\bar{f} \of \tilde{f}) &= p'_1 &       p'_2 \of (\bar{f} \of \tilde{f}) &= p'_2
    \end{alignat*}
    With the outer triangles we get:
    \begin{alignat*}{5}
      p_1  &\;\of&&\; f  &= p_1  &\qquad p_2  &\;\of&\; f  &= p_2 \\
      p'_1 &\;\of&&\; f' &= p'_1 &\qquad p'_2 &\;\of&\; f' &= p'_2
    \end{alignat*}
    Note that $f$ and $f'$ are \emph{unique} and the first equations hold for $f = 1_P$ and $f = \tilde{f} \of \bar{f}$, thus $\tilde{f} \of \bar{f} = 1_P$.
    The second equations holds for $f' = 1_{P'}$ and for $f' = \bar{f} \of \tilde{f}$, thus $\bar{f} \of \tilde{f} = 1_{P'}$.
    Thus $P$ and $P'$ are isomorphic. \qedhere
  \end{proof}
\end{answer}

@@

\begin{answer}
  \begin{proof}
    We first define the projection functions $p_1 \from V \oplus W \to V, p_2 \from V \oplus W \to W$:
    \begin{alignat*}{1}
      p_1(x,y) &:= x \\
      p_2(x,y) &:= y
    \end{alignat*}
    Now let $A$ be a vector space and $f_1 \from A \to V, f_2 \from A \to W$ be linear functions.
    We define $\bar{f}(a) := (f_1(a), f_(2))$. We clearly have
    \begin{alignat*}{2}
      p_1(\bar{f}(a)) &= p_1(f_1(a), f_2(a)) = f_1(a) \\
      p_2(\bar{f}(a)) &= p_2(f_1(a), f_2(a)) = f_2(a).
    \end{alignat*}
    Thus the diagramm
    \[ \xymatrix{
        & A \ar[ldd]_{f_1} \ar@{.>}[d]|{\bar{f}\vphantom{\bar{\bar{f}}}} \ar[rdd]^{f_2} & \\
        & {V \oplus W} \ar[ld]^{p_1} \ar[rd]_{p_2} & \\
        V & & W
    } \]
    commutes. Now we have to show that $\bar{f}$ is the only map such that the above diagramm commutes.
    Let $\tilde{f} \from A \to V \oplus W$ be a map such that $p_1 \of \tilde{f} = f_1 \land p_2 \of \tilde{f} = f_2$.
    \[
      \tilde{f}(a) = \bigl(p_1(\tilde{f}(a)), p_2(\tilde{f}(a))\bigr) = \bigl(p_1(\bar{f}(a)), p_2(\bar{f}(a))\bigr) = \bar{f}(a)\qedhere
    \]
  \end{proof}
\end{answer}

@@

\begin{exercise}[Equalizer of vector spaces]
  Let
  $\xymatrix{
      X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
  }$
  be a diagramm in $Vect_K$ for a field $K$. Show that $core(s-t)$ is the equalizer for this diagramm.
\end{exercise}

\begin{answer}
  \begin{proof}
    Let $i \from core(s-t) \to X$ be a inclusion function. Let
    $\xymatrix{
      A \ar[r]^f & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
    }$
    be a commuting diagramm. Note that $f A \subseteq core(s-t)$.
    Thus we can define $\bar{f} \from A \to core(s-t)$ with $\bar{f}(a) := f(a)$.
    We see that $\forall a \in A: i(\bar{f}(a))=\bar{f}(a)=f(a)$, thus $i \of \bar{f} = f$.
    Let $\tilde{f} \from A \to core(s-t)$ be another map such that $i \of \tilde{f} = f$.
    Now $\tilde{f}(a) = i(\tilde{f}(a)) = i(\bar{f}(a)) = \bar{f}(a)$.
  \end{proof}
\end{answer}

>>>>

    Lets reconcider the definition of the \emph{product}:
    For every object $A \in \cat{A}$ and arrows $f_1 \from A \to X, f_2 \from A \to Y$ we get that there exists a unique map
    $\bar{f} \from A \to P$, such that
    \[ \xymatrix{
        &A \ar[ldd]_{f_1} \ar@{.>}[d]|{\bar{f}\vphantom{\bar{\bar{f}}}}
        \ar[rdd]^{f_2}&       \\
                &P \ar[ld]^{p_1} \ar[rd]_{p_2}                  &       \\
        X       &                                               &Y
    } \]
    commutes.
    Now we can instanciate $A$ with $P'$, $f_1$ with $p_1$, $f_2$ with $p_2$. By analogous instanciations we get the commuting diagram
    \[ \xymatrix{
      & P \ar@/_/[ldd]_{p_1} \ar@/^/[rdd]^{p_2} & \\
      & P \ar[ld]^{p_1} \ar@/_/@{.>}[dd]|{\bar{f}} \ar@{.>}[u]|{f} \ar[rd]_{p_2} & \\
      X & & Y \\
      & P' \ar[lu]_{p_1'} \ar@/_/@{.>}[uu]|{\tilde{f}} \ar@{.>}[d]|{f'} \ar[ru]^{p'_2} & \\
      & P' \ar@/^/[luu]^{p'_1} \ar@/_/[ruu]_{p'_2} &
    } \]
    for exactly one $f, f', \tilde{f}, \bar{f}$. With the inner triangles we get:
    \begin{alignat*}{2}
      p_1  \of \tilde{f} &= p'_1 &\qquad p_2  \of \tilde{f} &= p'_2 \\
      p'_1 \of \bar{f}   &= p_1  &       p'_2 \of \bar{f}   &= p_2
    \end{alignat*}
    By substitution we get
    \begin{alignat*}{2}
      p_1   \of (\tilde{f} \of \bar{f}) &= p_1  &\qquad p_2  \of (\tilde{f} \of \bar{f}) &= p_2 \\
      p'_1  \of (\bar{f} \of \tilde{f}) &= p'_1 &       p'_2 \of (\bar{f} \of \tilde{f}) &= p'_2
    \end{alignat*}
    With the outer triangles we get:
    \begin{alignat*}{5}
      p_1  &\;\of&&\; f  &= p_1  &\qquad p_2  &\;\of&\; f  &= p_2 \\
      p'_1 &\;\of&&\; f' &= p'_1 &\qquad p'_2 &\;\of&\; f' &= p'_2
    \end{alignat*}
    Note that $f$ and $f'$ are \emph{unique} and the first equations hold for $f = 1_P$ and $f = \tilde{f} \of \bar{f}$, thus $\tilde{f} \of \bar{f} = 1_P$.
    The second equations holds for $f' = 1_{P'}$ and for $f' = \bar{f} \of \tilde{f}$, thus $\bar{f} \of \tilde{f} = 1_{P'}$.
    Thus $P$ and $P'$ are isomorphic. \qedhere
  \end{proof}
\end{answer}

@@

\begin{answer}
  \begin{proof}
    We first define the projection functions $p_1 \from V \oplus W \to V, p_2 \from V \oplus W \to W$:
    \begin{alignat*}{1}
      p_1(x,y) &:= x \\
      p_2(x,y) &:= y
    \end{alignat*}
    Now let $A$ be a vector space and $f_1 \from A \to V, f_2 \from A \to W$ be linear functions.
    We define $\bar{f}(a) := (f_1(a), f_(2))$. We clearly have
    \begin{alignat*}{2}
      p_1(\bar{f}(a)) &= p_1(f_1(a), f_2(a)) = f_1(a) \\
      p_2(\bar{f}(a)) &= p_2(f_1(a), f_2(a)) = f_2(a).
    \end{alignat*}
    Thus the diagram
    \[ \xymatrix{
        & A \ar[ldd]_{f_1} \ar@{.>}[d]|{\bar{f}\vphantom{\bar{\bar{f}}}} \ar[rdd]^{f_2} & \\
        & {V \oplus W} \ar[ld]^{p_1} \ar[rd]_{p_2} & \\
        V & & W
    } \]
    commutes. Now we have to show that $\bar{f}$ is the only map such that the above diagram commutes.
    Let $\tilde{f} \from A \to V \oplus W$ be a map such that $p_1 \of \tilde{f} = f_1 \land p_2 \of \tilde{f} = f_2$.
    \[
      \tilde{f}(a) = \bigl(p_1(\tilde{f}(a)), p_2(\tilde{f}(a))\bigr) = \bigl(p_1(\bar{f}(a)), p_2(\bar{f}(a))\bigr) = \bar{f}(a)\qedhere
    \]
  \end{proof}
\end{answer}

@@

\begin{exercise}[Equalizer of vector spaces]
  Let
  $\xymatrix{
      X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
  }$
  be a diagram in $Vect_K$ for a field $K$. Show that $core(s-t)$ is the equalizer for this diagram.
\end{exercise}

\begin{answer}
  \begin{proof}
    Let $i \from core(s-t) \to X$ be a inclusion function. Let
    $\xymatrix{
      A \ar[r]^f & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
    }$
    be a commuting diagram. Note that $f A \subseteq core(s-t)$.
    Thus we can define $\bar{f} \from A \to core(s-t)$ with $\bar{f}(a) := f(a)$.
    We see that $\forall a \in A: i(\bar{f}(a))=\bar{f}(a)=f(a)$, thus $i \of \bar{f} = f$.
    Let $\tilde{f} \from A \to core(s-t)$ be another map such that $i \of \tilde{f} = f$.
    Now $\tilde{f}(a) = i(\tilde{f}(a)) = i(\bar{f}(a)) = \bar{f}(a)$.
  \end{proof}
\end{answer}
